Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-08-06 Cost_manuscript.Rmd
editing the manuscript, and adding info on modeling

  Foraging theory posits that accepting a delayed reward should depend on the opportunity cost (OC) of time incurred in obtaining it, given by the richness of the environment (Stephens and Krebs, 1977). In this study, the richness of the environment was manipulated by the length of the handling and travel times (i.e. shorter travel times produced richer environments). Since time combinations were fixed per block, we calculated each block’s optimal accept/reject strategy by computing all decision strategies according to the following equation:

@@

  where *R*, *H*, and *p* are the reward, handling time, and acceptance probability of offer i, respectively, and T is the travel time. This gave us the total reward per second attainable in each block as a function of the lowest amount accepted (i.e. acceptance threshold). Figure 2 shows possible earnings for each choice strategy, as well as the lowest amount participants should accept in order to maximize their rewards (circled dots). For example, a participant in a 10 s handling block should accept 8-cent and 20-cent rewards (and reject 4-cent rewards) to maximize their reward rate. Note that accepting every offer was often detrimental to participant earnings. Significant deviations from the optimal strategy can be understood as reflecting changes in subjective OC. Beyond comparing among cost types, this approach allowed us to test if certain demands boosted the value of offers (i.e. over-accepting low offers).

@@

  All analyses were performed in R 3.5.2 (CITE). First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a mixed-effects logistic regression to predict trial-wise acceptances, using a random intercept to account for biases in participant behavior. We included regressors for handling time and reward, as well as two covariates that probed the influence of recent history on choices. The first was an autoregressive regressor containing the number of consecutive offer misses prior to a given trial (misses could be due to quits or lapses in performance that forced participants to travel). The second regressor tracked the sum of the previous *n* offers. We identified the *n* that minimized Akaike's Information Criterion (AIC) among 6 versions of the model (i.e. tracing from 2 previous trials up to 7, beyond which the model failed to converge), and reported all coefficients from the winning model.

  Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion accepted (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cost). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR).

@@

  In order to further look at the effect of handling, offer, and cost type, we computed the probability of accepting a trial with a mixed-effects logistic regression using the lme4 package (CITE). Based on the task structure and our main question, our a priori model of interest included cost condition, handling time, and reward amount as fixed main effects, and participant ID as a random intercept. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition. We ran three versions of the model with different reference conditions, in order to test all pairwise differences among the four cost conditions. (**exponentiate coefficients to make them more interpretable?**) We then examined whether this a priori model outperformed both simpler and more complex models. We used both AIC and Bayesian Information Criterion (BIC) to determine the model that minimized the negative log-likelihood while penalizing the addition of parameters. The regression with each combination of predictors was fitted in the following order: 1) intercept only; 2) condition only; 3) handling time only; 4) offer amount only; 5) handling time and offer amount (added to the preregistered comparison); 6) condition, handling time, and reward main effects (a priori model from above); 7) adding a handling-by-reward interaction; and 8) all possible two-way interactions. We predicted that model 6 (the a priori setup) would have the lowest AIC and BIC. Nested models with similar AIC were statistically compared using an analysis-of-deviance. The significance test was computed as the probability of the reduction in deviance, based on a chi-square distribution with degrees of freedom equal to the difference in the number of parameters between models.

  Finally, we modeled the subjective opportunity cost in each condition. We started by examining whether participants were confident and consistent in their choices. We visually examined survival curves indicating at what point during the handling time participants quit a trial (censored points included trial completions and forced travels), in addition to quitting time distributions. Next, we addressed the prediction that participants would display consistent choice patterns throughout the session (i.e. that the reward amounts they accepted in a given timing condition would be similar throughout the experiment). We computed each participant’s total proportion of acceptances pre- and post-midpoint for every block, and performed paired permutations (5000 iterations) to compare their mean proportion of acceptances. We then confirmed that the relative costs observed with from the whole experimental session were still present on each half of the experiment. The winning model from the mixed-effects logistic comparison was applied to each half separately. In each case, the reference cost category was rotated in order to assess pairwise differences among all costs. We evaluated the consistency of the relative costs by visually comparing the patterns of significant differences between experimental halves.

  Once we established choice stability, we proceeded to estimate the subjective OC of each type of demand (see Operationalization of Cost). OC was computed as the product of a free parameter (gamma) and the handling time. Gamma can be interpreted as the subjective value of each second spent in an environment. If a trial's reward rate (i.e. offer amount divided by the handling time) fell below gamma, a participant was better off finding more profitable alternatives. In other words, individuals with higher gammas required higher rewards in order to accept a given offer. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

@@

  Both gamma ($\gamma$) and the inverse temperature parameter ($\beta$) of the softmax function were estimated independently for each participant, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE). The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group. We then used a two-sided one-sample t-tests to compare each group's gammas to the average optimal earning rate (0.74, see Operationalization of Cost), allowing us to determine whether any demand boosted an offer's value. Finally, we determined which cost type produced the highest discounting by comparing gammas with an ANOVA (with condition as a factor), followed by post-hoc permutations (5000 iterations) comparing mean gammas per group. **ADD ADDITIONAL MODELS FROM LITERATURE**

@@

  The analytical pipeline was mostly preserved from the between-subject experiment. We first ran tests to determine whether decision makers integrated reward information. We performed a logistic regression per participant to predict trial-wise acceptances based on reward amounts. The resulting beta coefficients were pooled across all participants, and a one-sample rank-sum test was used to examine whether they were significantly positive or negative (compared to 0). Next, we explored the effects of choice history by adding an autoregressive covariate containing the number of consecutive recent quits. We then confirmed that the under- and over-acceptance were detrimental to total earnings by performing the same quadratic linear model explained above. To determine the optimality of the group’s decisions, we examined whether each cost type produced a bias to over or under-accept at 4 and 8 cents (assuming that 20 cents were always be accepted, given our design). The reward-maximizing strategy was to always reject 4 cents and accept 8 cents, yielding a combined optimal proportion of acceptances of 50% (see Operationalization of Costs for details). We performed a two-sided one-sample chi-squared test of proportions against the null probability of 0.5 for each type of cost. Therefore, a significant difference indicated that participants either over or under accepted rewards.

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and participant ID as a random intercept (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in the following order: 1) intercept only; 2) cost only; 3) reward only; 4) cost and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

@@

  **adapt once we decide on the model candidates** To estimate the subjective opportunity cost of each condition, we used a logistic function to model each participant’s probability of accepting a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. As before, OC was computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function were estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group were compared using a repeated measures ANOVA with cost as a factor. We performed pairwise paired permutations as post-hoc tests.

>>>>

  Foraging theory posits that accepting a delayed reward should depend on the opportunity cost (OC) of time incurred in obtaining it, given by the richness of the environment (Stephens and Krebs, 1977). In this study, the richness of the environment was manipulated by the length of the handling and travel times, where shorter travel times produced richer environments. Since time combinations were fixed per block, we calculated each block’s optimal accept/reject strategy by computing all decision strategies according to the following equation:

@@

  where *R*, *H*, and *p* are the reward, handling time, and acceptance probability of offer i, respectively, and T is the travel time. This gave us the total reward per second attainable in each block as a function of the acceptance threshold (i.e. the lowest amount accepted). Figure 2 shows possible earnings for each choice strategy, as well as the lowest amount participants should accept in order to maximize their rewards (circled dots). For example, a participant in a 10 s handling block should accept 8-cent and 20-cent rewards (and reject 4-cent rewards) to maximize their reward rate. Note that accepting every offer was often detrimental to participant earnings. Significant deviations from the optimal strategy can be understood as reflecting changes in subjective OC. This approach allowed us to test if certain demands boosted the value of offers, expressed as a subjective OC that fell below the optimal rate.

@@

  All analyses were performed in R 3.5.2 (CITE). First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a mixed-effects logistic regression to predict trial-wise acceptances, using a random intercept to account for biases in participant behavior. We included regressors for handling time and reward, as well as two covariates that probed the influence of recent history on choices. The first was an autoregressive regressor containing the number of consecutive offer misses prior to a given trial (misses could be due to quits or forced travels). The second regressor tracked the sum of the previous *n* offers. We identified the *n* that minimized Akaike's Information Criterion (AIC) among 6 versions of the model (tracing from 2 previous trials up to 7, beyond which the model failed to converge), and reported all coefficients from the winning model.

  Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion of trials completed (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, we performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cocitst). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR, CITE Y&..).

@@

  In order to further look at the effect of handling, offer, and cost type, we computed the probability of accepting a trial with a mixed-effects logistic regression, using the lme4 package (CITE). Based on the task structure and our main question, our a priori model of interest included cost condition, handling time, and reward amount as fixed main effects, and participant ID as a random intercept. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition. We ran three versions of the model with different reference conditions, in order to test all pairwise differences among the four cost conditions. (**exponentiate coefficients to make them more interpretable?**) We then examined whether this a priori model outperformed both simpler and more complex models. We used both AIC and Bayesian Information Criterion (BIC) to determine the model that minimized the negative log-likelihood while penalizing the addition of parameters. The regression with each combination of predictors was fitted in a forward order: 1) intercept only; 2) condition only; 3) handling time only; 4) offer amount only; 5) handling time and offer amount (added to the preregistered comparison); 6) condition, handling time, and reward main effects (a priori model from above); 7) adding a handling-by-reward interaction; and 8) all possible two-way interactions. We predicted that model 6 (the a priori setup) would have the lowest AIC and BIC. Nested models with similar AIC were statistically compared using an analysis-of-deviance. The significance test was computed as the probability of the reduction in deviance, based on a chi-square distribution with degrees of freedom equal to the difference in the number of parameters between models.

  Before modeling each individual's overarching subjective OC, we examined whether participants were consistent and stable in their choices. For consistency, we visually examined survival curves indicating at what point during the handling time participants quit a trial (censored points included trial completions and forced travels), in addition to quitting time distributions. Stability was assessed by computing each participant’s total proportion of acceptances pre- and post-midpoint for every block, comparing the mean proportion of acceptances using paired permutations (5000 iterations). We then tested whether the relative differences in costs observed using data from the whole experimental session were still present on each half of the experiment. The winning model from the mixed-effects logistic comparison was applied to each half separately. In each case, the reference cost category was rotated in order to assess pairwise differences among all costs. We evaluated the consistency of the relative costs by visually comparing the patterns of significant differences between experimental halves.

  Finally, we fitted the subjective OC of each type of demand (see Operationalization of Cost). OC was computed as the product of a free parameter (gamma) and the handling time. Gamma can be interpreted as the subjective value of each second spent in an environment. If a trial's reward rate (i.e. offer amount divided by the handling time) fell below gamma, a participant was better off finding more profitable alternatives. In other words, individuals with higher gammas required higher rewards in order to accept a given offer. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

@@

  Both gamma ($\gamma$) and the inverse temperature parameter ($\beta$) of the softmax function were estimated independently for each participant, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE **or grid search?**). The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group. We then used a two-sided one-sample t-tests to compare each group's gammas to the average optimal earning rate (0.74, see Operationalization of Cost), allowing us to determine whether any demand boosted an offer's value. Finally, we determined which cost type produced the highest discounting by comparing gammas with an ANOVA (with condition as a factor), followed by post-hoc permutations (5000 iterations) comparing mean gammas per group.

@@

  The analytical pipeline was mostly preserved from the between-subject experiment. We first ran tests to determine whether decision makers integrated reward information. **We performed a logistic regression per participant to predict trial-wise acceptances based on reward amounts. The resulting beta coefficients were pooled across all participants, and a one-sample rank-sum test was used to examine whether they were significantly positive or negative (compared to 0). Next, we explored the effects of choice history by adding an autoregressive covariate containing the number of consecutive recent quits (mixed-effects)**. We then confirmed that the under- and over-acceptance were detrimental to total earnings by performing the same quadratic linear model explained above. To determine the optimality of the group’s decisions, we examined whether each cost type produced a bias to over or under-accept at 4 and 8 cents (assuming that 20 cents were always be accepted, given our design). The reward-maximizing strategy was to always reject 4 cents and accept 8 cents, yielding a combined optimal proportion of acceptances of 50% (see Operationalization of Costs for details). We performed a two-sided one-sample chi-squared test of proportions against the null probability of 0.5 for each type of cost. Therefore, a significant difference indicated that participants either over or under accepted rewards.

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and participant ID as a random intercept (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in a forward order: 1) intercept only; 2) cost only; 3) reward only; 4) cost and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

@@

  **old** To estimate the subjective opportunity cost of each condition, we used a logistic function to model each participant’s probability of accepting a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. As before, OC was computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function were estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group were compared using a repeated measures ANOVA with cost as a factor. We performed pairwise paired permutations as post-hoc tests.

**new** The adaptive nature of choices in this experiment demanded more flexibility than the single-parameter model fitted above. Based on the results, an adequate model of behavior should meet X criteria: 1) an initial preference for cognitive effort trials 2) that this preference disappeared as acceptances devolved to a common rate across costs; 3) show that participants kept track of recent offer history; 4) that these effects were robust to block order effects; 5) .... X) be applicable to both experiments
